\section{Conclusion}\label{sec:conclusion}

We went into detail about the response functions in correlated financial
markets. We define the trade time scale and physical time scale to compute the
self- and cross-response functions for six companies with the largest average
market capitalization for three different economic sectors of the S\&P index
in 2008.
Due to the characteristics of the data used, we had to classify and sampling
values to obtain the corresponding quantities in different time scales.
The classification and sampling of the data had impact on the results, making
them smoother or stronger, but always keeping their shape and behavior.

The response functions were analyzed according to the time scales. To compare
price response functions from different scales, we used the same midpoint
prices in physical time scale with the corresponding trade signs in trade time
scale or physical time scale. This assumption allowed us to get an idea of how
representative was the behavior obtained in both cases. For trade time scale,
the signal is weaker due to the large averaging values from all the trades in a
year. In the physical time scale, the response functions had less noise and
their signal were stronger. The activity in every second highly impact the
responses. As the response functions can not grow indefinitely with the time
lag, they increase to a peak, to then decrease. It can be seen that the market
needs time to react and revert the growing. In both time scale cases depending
on the stocks, two characteristics behavior were shown. In one, the time lag
was large enough to show the complete increase-decrease behavior. In the other
case, the time lag was not enough, so some stocks only showed the growing
behavior.

We modify the response function to add a time shift parameter. With this
parameter we wanted to analyze the importance in the order of the relation
between returns and trade signs. In trade time scale and physical time scale we
found similar results. When we shift the order between returns and trade signs,
the information from the relation between them is temporarily lost and as
outcome the signal does not have any meaningful information. When the order is
recovered, the response function grows again, showing the expected shape.
We showed that this is not an isolated conduct, and that all the shares used in
our analysis exhibit the same behavior. Thus, even if they are values of time
shift that can give a response function signal, from the theory this time shift
should be a value between $t_{s} = [0,2]$.

We analyzed the impact of the time lag in the response functions. We divided
the time lag in a short and long time lag. The response function that depended
on the short time lag, showed a stronger response. The long response function
depending on the stock could take negative and non-negative values. However, in
general the influence were not intense.

Finally, we checked the spread impact in self-response functions. We divided
530 stocks from the NASDAQ stock market in five groups depending on the year
average spread of every stock. The response functions signal were stronger for
the group of stocks with the larger spreads and weaker for the group of stocks
with the smaller spreads.